---
title: "Practical Data Analysis Homework 4"
author: "Christopher Costello"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.path = "my_report_files/"
  )
```


# Question 2

How many rows are available in the Measurements table of the Smith College Wideband Auditory Immittance database?

```{r question 2}
library(tidyverse)
library(RMariaDB)
con <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)
Measurements <- tbl(con, "Measurements")

dbGetQuery(con,
           "
           SELECT SUM(1) AS num_rows
           FROM `Measurements`
           ")

# Disconnect database
```

From this query, we find that the number of rows in the Measurements table is 5,052,304.

# Question 3

Identify what years of data are available in the flights table of the airlines database.

```{r question 3}
library(tidyverse)
library(mdsr)
library(RMariaDB)
con <- dbConnect_scidb("airlines")

dbListTables(con)

tbl(con, "flights")

dbGetQuery(con,
           "
           SELECT DISTINCT year
           FROM flights
           ")
```

# Question 4

Use the dbConnect_scidb function to connect to the airlines database to answer the following problem.

The `airlines` database is already connected from Question 3.

## Question 4.1

How many domestic flights flew into Dallas-Fort Worth (DFW) on May 14, 2015?

```{r question 4.1}
dbGetQuery(con,
          "
          SELECT SUM(1) AS number_of_flights
          FROM flights
          WHERE
            dest = 'DFW' AND
            year = 2015 AND
            month = 5 AND
            day = 14
          ")
```

## Question 4.2

Of all the destinations from Chicago O’Hare (ORD), which were the most common in 2015?

```{r question 4.2}
dbGetQuery(con,
          "
          SELECT
            dest,
            SUM(1) AS number_of_flights
          FROM
            flights
          WHERE
            origin = 'ORD' AND
            year = 2015
          GROUP BY
            dest
          ORDER BY
            number_of_flights DESC
          LIMIT 10
          ")
```

## Question 4.3

Which airport had the highest average arrival delay time in 2015?

```{r question 4.3}
dbGetQuery(con,
          "
          SELECT
            dest,
            avg(arr_delay) AS average_arrival_delay
          FROM flights
          WHERE
            year = 2015
          GROUP BY
            dest
          ORDER BY
            average_arrival_delay DESC
          LIMIT 1
          ")
```